\textsc{Beyond the Iceberg Hypothesis: \\Opening the Black Box of Transport Costs}

### 2017-05-19 redaction/dhp_sept2016.tex
changement nouvelle methode

Figure \ref{fig:totalTC_compeffects_excl} reports the results. Panel (a) displays the time changes in the overall transport costs, by transport mode. In Panels (b) and (c), we report the time changes of the ad-valorem costs and the additive costs respectively (re-scaled to start from the reference value 100), for each transport mode.

@@

Consider first the time trends in transport costs at the aggregate level (Figure \ref{fig:totalTC_compeffects_excl}). Two main results emerge. First, we do not find evidence of a significant difference between the ``raw'' transport costs and the `pure'' transport costs in Air transport. This result for each of the two components, additive and ad-valorem (Figure \ref{fig:totalTC_compeffects_excl}, left panels). Air transport costs were reduced by 50\% between 1974 and 2013, and this is mainly attributable to a reduction in the ``pure'' transport costs. This stands in contrast with \citet{hummels2007}, which obtained that the ``pure'' transport costs had decreased more (over 1973- 2004) than the unfitted ones, suggesting an important role to composition effects. As noted above, this difference of results can be explained by the difference in methodology. Whereas \citet{hummels2007} attributes all changes that are not changes in the ad-valorem components to composition effects, we show that part (and as a matter of fact, most) of these changes come from a reduction in the ``pure'' transport costs, in their additive dimension.

>>>>

Figure \ref{fig:totalTC_compeffects_excl} reports the results. In Panels (a) and (b), we report the time changes of the ad-valorem costs and the additive costs respectively. Panel (c) displays the time changes in the total transport costs (built as the sum of the two components as detailed above). In the three cases, the series are reported as indices, starting from the reference value 100 in 1974 and by transport mode.

@@

Consider first the time trends in transport costs at the aggregate level (Figure \ref{fig:totalTC_compeffects_excl}). Three main results emerge. First, we do not find evidence of a significant difference between the raw transport costs and the ``pure'' transport costs in Air transport, for both the additive and ad-valorem components (Figure \ref{fig:totalTC_compeffects_excl}, left panels). Air transport costs were reduced by 50\% between 1974 and 2013, and this is mainly attributable to a reduction in the ``pure'' transport costs. This stands in contrast with \citet{hummels2007}, which obtained that the ``pure'' transport costs had decreased more (over 1973- 2004) than the unfitted ones, suggesting an important role to composition effects. As noted above, this difference of results can be explained by the difference in methodology. Whereas \citet{hummels2007} attributes all changes that are not in the ad-valorem component to composition effects, we show that part (and as a matter of fact, most) of these changes do come from a reduction in the ``pure'' transport costs, in their additive dimension.

Second, composition effects are more pronounced in ocean transport. Considering the raw series, maritime transport costs have decreased by 60\% over the period, which can be decomposed in a 50\% decrease in transport costs \textit{per se}, and a 10\% reduction that comes from composition effects. Precisely, the role of composition effects is significant for the multiplicative component, whereas the reduction in the additive costs is mostly attributable to changes in ``pure'' additive costs. In line with \citet{hummels2007}, these results confirm the important role of composition effects in maritime transport costs. The direction of the effects differs though. Whereas  \citet{hummels2007} obtains a more pronounced reduction in the ``pure'' transport costs than the unfitted series, we obtain the opposite result. On top of the reduction of the ``pure'' transport costs, overall maritime transport costs have reduced over time because of the composition of flows, \textbf{one potential explanation being the reduction in the share of bulky goods linked to the declining share of primary goods in total trade. Aller plus loin? mettre la part du primary dans le temps? Pertinence de cette hypothese?}
